Conflict-Free Vertex Coloring Of Planar Graphs

### 2017-05-01 paper/paper.tex
Camera ready... hopefully

The \emph{conflict-free coloring} problem is a relaxed variation of the vertex coloring problem. The conflict-free coloring problem does not aim to color \emph{every} vertex such that it is not connected to a vertex of the same color. Rather, it aims to color \emph{some} vertices such that the neighborhood of every vertex contains at least one unqiuely colored vertex.

@@

The neighborhood of a vertex, denoted $N_G(v)$, is a set of all vertices adjacent to $v$. A \emph{closed} neighborhood consists of the vertices adjacent to $v$ plus $v$ itself. For example, let's examine $N_{G_1}(1)$ from Figure \ref{fig:graph}. The open neighborhood of 1 is $\{2, 5\}$ while the closed neighborhood is $\{1, 2, 5\}$. For the rest of the paper, the term \emph{neighborhood} will refer to closed neighborhoods unless stated otherwise.

@@

The distance of a path can be thought of as the smallest number of edges needed to get from one vertex to another vertex. For example, in Figure \ref{fig:graph}, the distance from vertex 3 to vertex 4 would be 3 as it takes 3 edges to get there. A \emph{distance-3-set} is a set of vertices that all have exactly distance 3 from each other. An example of a distance-3-set from Figure \ref{fig:graph} would be $\{3,4\}$. \cite{bondy1976graph,west2001introduction}

@@

\subsection{Graph Coloring}
\label{sec:coloring}
A \emph{vertex coloring} is an assignment of colors to each vertex of a graph $G$. A \emph{proper vertex coloring} assigns colors such that no two adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors, i.e. $|S|$, needed to properly color $G$. The \emph{vertex coloring problem} (VCP), when given a simple graph $G$, is to find $\chi(G)$. \cite{bondy1976graph}

A graph $G$ is said to be \emph{k-colorable} if it can be colored using $k$ or fewer colors, i.e. $\chi(G) \leq k$. A graph having $\chi(G) = k$ is said to be a \emph{k-chromatic} graph. The \emph{k-colorability} problem asks if a graph can be colored using $k$ colors. This problem is slightly different than the VCP as it is decision problem rather than an optimization problem. An example of a 3-colorable graph and a vertex coloring is shown in Figure \ref{fig:vcp-example}.

@@

A \emph{conflict-free k-coloring} of a simple graph $G$ assigns colors, $\{1, 2, \ldots, k\}$, to a subset $P \subseteq V$ of vertices such that $\forall v \in V$, there is a vertex $u \in N(v)$ where the color of $u$ is unique in the neighborhood of $v$. The set $V \setminus P$ contains any uncolored vertices. Generally, a \emph{conflict-free coloring} of a graph assigns colors to some of the vertices such that, for every vertex $v$, there is a unique color assigned to a vertex among $v$ and $v$'s neighbors. The vertex $u$ can be thought of the \emph{conflict-free neighbor} of $v$. The \emph{conflict-free chromatic number} of G, denoted $\chi_{CF}(G)$, is the smallest $k$ for which a conflict-free coloring exists. The \emph{conflict-free coloring problem} (CFCP) aims to find $\chi_{CF}(G)$. \cite{abel2017three}

@@

A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $G_1$ where $\gamma(G_1) = 2$. The \emph{k-conflict-free dominating set} problem asks if a given k-coloring of a graph colors the minimum amount of vertices needed to have a coloring that utilizes $k$ colors.

@@

The k-colorability problem as well as the conflict-free k-colorability problem are NP-complete. This can be proven by using a reduction from a known NP-complete problem \cite{garey2002computers}. If we find a known hard problem $Y$, we can prove that another problem $X$ is hard by a reduction from $Y$ to $X$. The VCP and the CFCP have both been shown to be NP-hard \cite{abel2017three,moret1998theory}. The \emph{k-colorability} problem is proven to be NP-complete with a reduction from 3-SAT, another well-known NP-complete problem \cite{sharma2012new}. The \emph{conflict-free k-colorability problem} is shown to be NP-complete by a reduction from k-colorability \cite{abel2017three}.

It is often desirable and necessary to use approximation algorithms, known as heuristics, to solve complex problems like the VCP and the CFCP. In our case, heuristics are polynomial-time algorithms that give good, but not optimal, solutions to our optimization problems. A polynomial-time approximation scheme (PTAS) is any polynomial-time approximation algorithm that takes an instance of an optimization problem and a parameter $\epsilon > 0$ and produces a solution that is within a factor $(1 + \epsilon)$ (for minimization problems) of being optimal.

@@

Imagine we have a set of readers, $R$. Suppose they all are assigned the same frequency. We would like to schedule each reader $r \in R$ a time slot $t(r)$ where the reader $r$ will be active. We have a set of tags $P$ (i.e. automobile RFID chips). We can say that $P$ is read by our schedule if for every tag $p \in P$, there is at least one reader $r \in R$ and a time $t$ such that $p$ is read by $r$ at time $t$. We want to minimize the total time slots used by the schedule. This can be accomplished by modeling the situation as a conflict-free coloring problem. Again, we aim to find the minimum number of colors needed. \cite{cheilaris2014strong, smorodinsky2013conflict}

@@

The last step of the algorithm is to remove all isolated paths from $G$. Since we are left with only two vertices with no edges connecting them, we have 2 isolated paths. We remove these isolated paths from $G$ and are left with an empty graph. Since the graph is empty, we exit the outer $while$ loop and execute line 17. We color all removed isolated paths using color $i$, which is color 2 (say blue). Since we had 2 isolated paths, which are really just two vertices 5 and 7, we color each vertex with color 2.

@@

@@

There are many applications of the CFCP that benefit from the recent research into conflict-free coloring. It is easier to minimize frequency assignment for wireless networks and design RFID networks. Current research focuses on finding bounds and properties on specific cases such as planar graphs, interval graphs, types of hypergraphs, and more.

There is also recent interest in variations of the CFCP such as requiring \emph{another} vertex (not the selected vertex) to be colored within the neighborhood of a selected vertex. This allows for applications such as guiding a specific robot to other locations (i.e. the robot and its destination would both be a distinct unique color). Overall, finding properties and heuristics for conflict-free coloring is beneficial and required for efficiently solving real-world problems.

>>>>

The \emph{conflict-free coloring} problem is a relaxed variation of the vertex coloring problem. The conflict-free coloring problem does not aim to color \emph{every} vertex. Rather, it aims to color \emph{some} vertices such that the neighborhood of every vertex contains at least one uniquely colored vertex.

@@

The open neighborhood of a vertex is a set of all vertices adjacent to $v$. A \emph{closed} neighborhood, denoted $N_G(v)$, consists of the vertices adjacent to $v$ plus $v$ itself. For example, let's examine $N_{G_1}(1)$ from Figure \ref{fig:graph}. The open neighborhood of 1 is $\{2, 5\}$ while the closed neighborhood is $\{1, 2, 5\}$. For the rest of the paper, the term \emph{neighborhood} will refer to closed neighborhoods unless stated otherwise.

@@

The distance of a path can be thought of as the smallest number of edges needed to get from one vertex to another vertex. For example, in Figure \ref{fig:graph}, the distance from vertex 3 to vertex 4 would be 3 as it takes 3 edges to get there. A \emph{distance-3-set} is a set of vertices that all have exactly pairwise distance 3 from each other. An example of a distance-3-set from Figure \ref{fig:graph} would be $\{3,4\}$. \cite{bondy1976graph,west2001introduction}

@@

\subsection{Graph Coloring}
\label{sec:coloring}
A \emph{vertex coloring} is an assignment of colors to each vertex of a graph $G$. A \emph{proper vertex coloring} assigns colors such that no two adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors needed to properly color $G$. The \emph{vertex coloring problem} (VCP), when given a simple graph $G$, is to find $\chi(G)$. \cite{bondy1976graph}

A graph $G$ is said to be \emph{k-colorable} if it can be colored using $k$ or fewer colors, i.e. $\chi(G) \leq k$. A graph having $\chi(G) = k$ is said to be a \emph{k-chromatic} graph. The \emph{k-colorability} problem asks if a graph can be colored using $k$ colors. This problem is slightly different than the VCP as it is a decision problem rather than an optimization problem. An example of a 3-colorable graph and a vertex coloring is shown in Figure \ref{fig:vcp-example}.

@@

A \emph{conflict-free k-coloring} of a simple graph $G$ assigns colors, $\{1, 2, \ldots, k\}$, to a subset $P \subseteq V$ of vertices such that $\forall v \in V$, there is a vertex $u \in N(v)$ where the color of $u$ is unique in the neighborhood of $v$. The set $V \setminus P$ is the uncolored vertices, if any. Generally, a \emph{conflict-free coloring} of a graph assigns colors to some of the vertices such that, for every vertex $v$, there is a unique color assigned to a vertex among $v$ and $v$'s neighbors. The vertex $u$ can be thought of the \emph{conflict-free neighbor} of $v$. The \emph{conflict-free chromatic number} of G, denoted $\chi_{CF}(G)$, is the smallest $k$ for which a conflict-free coloring exists. The \emph{conflict-free coloring problem} (CFCP) aims to find $\chi_{CF}(G)$. \cite{abel2017three}

@@

A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The set of colored vertices in a conflict-free coloring is a dominating set \cite{abel2017three}. The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. An example of a dominating set is $\{3, 5\}$ of graph $G_1$ where $\gamma(G_1) = 2$. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. The \emph{k-conflict-free dominating set} problem asks if a given k-coloring of a graph colors the minimum amount of vertices needed to have a conflict-free coloring that utilizes $k$ colors.

@@

The k-colorability problem as well as the conflict-free k-colorability problem are NP-complete. This can be proven by using a reduction from a known NP-complete problem \cite{garey2002computers}. If we find a known hard problem $Y$, we can prove that another problem $X$ is hard by a reduction from $Y$ to $X$. The VCP and the CFCP have both been shown to be NP-hard \cite{abel2017three,moret1998theory}. The k-colorability problem is proven to be NP-complete with a reduction from 3-SAT, another well-known NP-complete problem \cite{sharma2012new}. The conflict-free k-colorability problem is shown to be NP-complete by a reduction from k-colorability \cite{abel2017three}.

It is often desirable and necessary to use approximation algorithms, called heuristics, to solve complex problems like the VCP and the CFCP. In our case, heuristics are polynomial-time algorithms that give good, but not optimal, solutions to our optimization problems. A polynomial-time approximation scheme (PTAS) is any polynomial-time approximation algorithm that takes an instance of an optimization problem and a parameter $\epsilon > 0$ and produces a solution that is within a factor $(1 + \epsilon)$ (for minimization problems) of being optimal.

\vspace{-0.2cm}

@@

Imagine we have a set of readers, $R$. Suppose they all are assigned the same frequency. We would like to schedule each reader $r \in R$ a time slot $t(r)$ where the reader $r$ will be active. We have a set of tags $P$ (i.e. product RFID chips). We can say that $P$ is read by our schedule if for every tag $p \in P$, there is at least one reader $r \in R$ and a time $t$ such that $p$ is read by $r$ at time $t$. We want to minimize the total time slots used by the schedule. This can be accomplished by modeling the situation as a conflict-free coloring problem. Again, we aim to find the minimum number of colors needed. \cite{cheilaris2014strong, smorodinsky2013conflict}

@@

The last step of the algorithm is to remove all isolated paths from $G$. Since we are left with only two isolated vertices, which are trivially isolated paths, we have 2 isolated paths. We remove these isolated paths from $G$ and are left with an empty graph. Since the graph is empty, we exit the outer $while$ loop and execute line 17. We color all removed isolated paths using color $i$, which is color 2 (say blue). Since we had 2 isolated paths, which are really just two vertices 5 and 7, we color each vertex with color 2.

@@

\vspace{-0.15cm}

@@

There are many applications of the CFCP that benefit from the recent research into conflict-free coloring. It is easier to minimize frequency assignment for wireless networks and design RFID networks. Current research focuses on finding bounds and properties on specific cases such as interval graphs, types of hypergraphs, and more \cite{cheilaris2014strong,smorodinsky2013conflict}.

There is also recent interest in variations of the CFCP such as requiring \emph{another} vertex (not the selected vertex) to be colored within the neighborhood of a selected vertex. This allows for applications such as guiding a specific robot to other locations (i.e. the robot and its destination would both be a distinct unique color). Relaxing the proper vertex coloring problem to satisfy new requirements leads to new heuristics that can effectively solve more real-world problems.

\vspace{-0.15cm}
